\documentclass[pdftex,12pt,a4paper]{article}

\input{/home/boris/Dropbox/Public/tex_general/title_bor_utf8}

\emergencystretch=2em \voffset=-2cm \hoffset=-2cm
\unitlength=0.6mm \textwidth=18cm \textheight=26cm


\begin{document}
\parindent=0 pt % отступ равен 0

Final exam retake on mathematics for the Master degree students, ICEF
February, 3rd, 2012


Section A.  10 points for each problem.

\begin{enumerate}
\item Solve the bounded control problem with the free end value where $T>1$ is specified in advance but $x(T)$ is not.
\begin{equation}
\int_0^T(x-u)\,dt\to \max,\, x(0)=1 \nonumber
\end{equation}
$x'=u$, $0\leq u\leq x$. What guarantees that a maximizer would be found?

\item Two stochastic processes are defined by the system of SDE with the Brownian motions $W_{1t}$, $W_{2t}$ independent of each other
\begin{equation}
\begin{cases}
dX_t=(2+5t+X_t)dt+3dW_{1t} \\
dY_t=4Y_t dt+8Y_t dW_{1t}+6dW_{2t} \nonumber
\end{cases}
\end{equation}
Calculate $d(X_t Y_t)$

\item Find the value of the constant $a$ such that the process $X_t=W_t^3-a\int_0^t W_s\,ds$ is a martingale.

\item The process $Y_t$ is given by $Y_t=2W_t+5t$. The stopping time $\tau$ is given by $\tau=\min\{t|Y_t^2=100\}$. Find the distribution of the random variable $Y_\tau$ and the expected value $\E(\tau)$.


Hint: you may find the martingales $a^{Y_t}$ and $Y_t-f(t)$ useful

\item What is the expected value and variance of $W_t^2$ for $t>s$ given that $W_s=x$?

\item In the framework of Black and Scholes model find the price at the time $0$ of an asset which gives you the payoff $\max\{\ln(S_t),0\}$ at the time $t$. Here $S_t$ is the price of the underlying asset.
\end{enumerate}


Section B. 20 points for each problem. 

\begin{enumerate}
\item The goal of this exercise is to solve the SDE
\begin{equation}
dX_t=\frac{1}{X_t}dt+X_tdW_t \nonumber
\end{equation}
with initial condition  $X_0=1$.
\begin{enumerate}
\item Apply the Ito's lemma to the process $Y_t=\exp(f(t)+g(W_t))X_t^2$
\item Find non-constant functions $f$ and $g$ such that the coefficient before $dW_t$ in the expression for $dY_t$ is zero.
\item Find $X_t$. The final expression may contain a Riemann integral of some stochastic process.
\end{enumerate}

\item A typical firm with the production function $x=f(l)$ in an economy employs 1 unit of the capital paying for it $\bar{r}$. There are $n$ identical firms in this economy where $n(t)$ is a function of time. Let the output of a firm be $x(t)$. Then the aggregate supply equals $nx$. Assuming that the market is in the equilibrium we use the inverse demand function $p(nx)$, where $p'(y)<0$. Equation defining the dynamics of labor is given by $\frac{dl}{dt}=a(w-\bar{w})$, where $a>0$  and $\bar{w}$  is some equilibrium wage rate. The growth (or fall) in number of firms is governed by the equation $\frac{dn}{dt}=b(px-\bar{w}l-\bar{r})$, where $b>0$. Write down the system of ODE for the unknowns $l(t)$, $n(t)$, using the first-order condition for the profit-maximizing firm, find and classify the steady-state solutions (if any exist). Production function is twice differentiable and concave everywhere.

\end{enumerate}


\begin{comment}
The stochastic process is defined by the formula (geometric Brownian motion)
\begin{equation}
S_t=S_0\exp(\sigma W_t+(\alpha-\frac{1}{2}\sigma^2)t)
\end{equation}
 ,
            where  are nonrandom constants and  is a Brownian motion. 
(points) Consider a partition of the time line segment  consisting of   subintervals of the type  with the corresponding norm . Find the value of the limit .
(points ) Let . Prove that  is a martingale.


\end{comment}

\end{document}