\documentclass[pdftex,12pt,a4paper]{article}

\input{/home/boris/Dropbox/Public/tex_general/title_bor_utf8}

\emergencystretch=2em \voffset=-2cm \hoffset=-2cm
\unitlength=0.6mm \textwidth=18cm \textheight=26cm


\begin{document}
\parindent=0 pt % отступ равен 0
Exam 12.01.2011 on stochastic calculus. \\


Notation: $W_t$ is the standard Wiener process. \\


Part A (10 points each problem). Time allowed: 120 minutes.  \\

\begin{enumerate}
\item The joint probability density function of $X$ and $Y$ is given by
\begin{equation}
f(x,y)=\begin{cases}
x+y, \quad x\in[0;1],\, y\in [0;1] \\
0, \quad \mbox{otherwise}
\end{cases} \nonumber
\end{equation}
Find $\E(Y|X)$ in terms of $X$, find the probability density function of  $\E(Y|X)$
\item Consider the process $X_t=\int_0^t sW_s\,dW_s$. Find $\E(X_t)$, $Var(X_t)$, $Cov(X_t,W_t)$
\item The process $Y_t$ is given by $Y_t=2W_t+5t$. The stopping time $\tau$ is given by $\tau=\min\{t|Y_t^2=100\}$. Find the distribution of the random variable $Y_\tau$ and the expected value $\E(\tau)$.


Hint: you may find the martingales $a^{Y_t}$ and $Y_t-f(t)$ useful
\item Find $\P(W_2-W_1>2)$
\item In the framework of Black and Scholes model find the price of an asset which gives you the payoff of $1$ rubble only if the final price $S_t$ is at least two times bigger than the initial price $S_0$ of the asset. 
\item Consider the free end problem, where $T>0$ is not given
\begin{equation}
\int_0^T (\dot{x}^2-x+1)dt \to extr \nonumber
\end{equation}
At the left end $x(0)=0$
Find the optimal $T$ value and the extremal. Check that you solved the optimality problem or show the opposite.
\end{enumerate}

\newpage

Part B (20 points each prolem). Time allowed: 60 minutes.  \\

\begin{enumerate}
\item Consider the stochastic differential equation 
\begin{equation}
dX_t=(\sqrt{1+X_t^2}+0.5X_t)dt+\sqrt{1+X_t^2}dW_t,\quad X_0=0 \nonumber
\end{equation}
\begin{enumerate}
\item Suppose that $Y_t$ is another process that depens only on $X_t$, i.e. $Y_t=f(X_t)$. Find $dY$ using the Ito's lemma.
\item Find such function $f$ that the term before $dW$ in $dY$ is constant.
\item Find $X_t$
\item Sketch $\P(X_t>0)$ as the function of $t$.
\end{enumerate}

\item Consider the neoclassical optimal growth model
\begin{equation}
\int_0^{\infty}e^{-rt}\left( \bar{U}-\frac{1}{c(t)}\right)\,dt \to \max \nonumber
\end{equation}
subject to $\dot{k}=A\ln(1+k)-c-\delta k$, where $A>r>\delta>0$, $k(0)=k_0$, $\bar{U}>0$.
\begin{enumerate}
\item Derive necessary conditions, using the current value Hamiltonian
\item Sketch the phase diagram for this problem with the capital intensity and consumption labeled on the horizontal and vertival axes, respectively
\item Check that the steady state solution exists. Provide explanation.
\item Explore the stability of the steady state, using the Jacobian
\item Why are you sure the found growth path maximizes the discounted stream of utility?
\end{enumerate}
\end{enumerate}


\end{document}