\documentclass[pdftex,12pt,a4paper]{article}

\input{/home/boris/Dropbox/Public/tex_general/title_bor_utf8}

%\usepackage{showkeys} % показывать метки

\input{/home/boris/Dropbox/Public/tex_general/prob_and_sol_utf8}

%\title{Задачи по элементарной теории вероятностей и матстатистике}
%\author{Составитель: Борис Демешев, boris.demeshev@gmail.com}
%\date{\today}

\begin{document}

%\pagestyle{myheadings} \markboth{ТВИМС-задачник. Демешев Борис. roah@yandex.ru }{ТВИМС-задачник. Демешев Борис. roah@yandex.ru }
%\maketitle
%\tableofcontents{}

%\parindent=0 pt % отступ равен 0


Problem 1.

$ Y\sim U[0;1] $, $ X=Y^{3} $, $ A=\{Y>0.8\} $, $ Z=1_{A}-Y $. Find $ E(Z|X) $, $ E(X|Z) $, $ E(X|1_{A}) $

Solution:

We note that $ X=Y^{3} $ and $ Z=1_{A}-Y $ are deterministic functions of $ Y $, so:

\[ E(X|Z)=X \]

(+3 points)

and

\[ E(Z|X)=Z \]

(+3 points)


The most difficult:

\[E(X|1_{A})=E(Y^{3}|A)1_{A}+E(Y^{3}|A^{c})1_{A^{c}}\]

(+4 points)


\newpage
Problem 2.

$ Y_{t}=W_{t}^{3}+f(t)W_{t} $. Find $ f(t) $ if it is known that $ Y_{t} $ is a martingale.

Solution:

Using the Ito's lemma we find $ dY_{t}=3W_{t}^{2}dW_{t}+f'(t)W_{t}dt+ 3W_{t}dt$.

(+5 points for the use of the Ito's lemma)

$ Y_{t} $ is a martingale, so $ f'=-3 $

We find $ f'(t)=-3 $ and $ f(t)=-3t+c $ where $ c\in \mathbb{R} $

(+5 points for the rest. The most common error: missing $c$ gives -2 points penalty)


Alternative way:

Using the definition of a martingale:

\[E(Y_{t}|\mathcal{F}_{s})=Y_{s}\]

Thus we have an equation $ f(t)-f(s)=-3(t-s) $ for all $s$ and $t$

(+5 points)

Plugging $s=0$ we have $ f(t)=-3t+f(0)$.

(+5 points)


\newpage

Problem 3. 

\[ Y_{n}=\frac{nX_{1}+X_{2}+...+X_{n+1}}{2n} \]. In what sense does the $ Y_{n} $ converge? What is the limit? Find the expected value and the variance of the limit.

Solution:

A little bit of algebra and we have:

\[ Y_{n}=\frac{X_{1}}{2}+0.5\frac{X_{2}+...+X_{n+1}}{n} \]

The Strong Law of Large Numbers (as):

\[ \frac{X_{2}+...+X_{n+1}}{n} \to E(X_{i})=2010 \]

So (as):

\[ Y_{n} \to \frac{X_{1}}{2}+1005 \]

(+4 points for the correct limit)

And $ E(\lim Y_{n} )=2010$, $ Var(\lim Y_{n} )=2011/4$

(+1 point for the expected value, +2 points for the variance)

Almost surely convergence implies convergence in probability and in distribution.

(+1 point for all convergence types except $ L^{2} $)

We check $ L^{2} $ convergence:

\[E((Y_{n}-\frac{X_{1}+2010}{2})^{2})=\frac{1}{4}E((\bar{X}-2010)^{2})=\frac{1}{4}Var(\bar{X})=\frac{2011}{4n}\]

This tends to zero and we have established $ L^{2} $ convergence.

$ L^{2} $ convergence implies $ L^{1} $ convergence.

(+2 points for the $ L^{2} $ convergence)


\newpage

Problem B1.

\[ dY=\frac{W}{Y}dW-\frac{W^{2}}{2Y^{3}}dt \]

a) Verify whether $ Z_{t}=Y_{t}^{2} $ is a martingale

b) Solve this equation if $ Y_{0}=1 $

Solution:

a) Apply the Ito's lemma for $ Z_{t}=Y_{t}^{2} $:

\[ dZ=2YdY+\frac{1}{2}2(dY)^{2} \]

Where $ (dY)^{2}=\frac{W^{2}}{Y^{2}}dt $

(+5 points)

This simplifies to 

\[ dZ=WdW \]

There is no $ dt $ term and we conclude that $Z_{t}$ is a martingale.

(+5 points)

b) Calculation of the stochastic integral $ \int_{0}^{t}W_{s}dW_{s}=W_{t}^{2}-t $

(+5 points)

So we obtain  $ Z_{t}=Z_{0}+W_{t}^{2}-t $. And using initial conditions:

$ Y_{t}=\sqrt{1+W_{t}^{2}-t} $

(+5 points)







\bibliography{/home/boris/Dropbox/Public/tex_general/opit}
\printindex % печать предметного указателя здесь

\end{document}