\documentclass[12pt, a4paper]{article}

\usepackage[top=1.5cm, left=2cm, right=2cm, bottom=1.5cm]{geometry} % размер текста на странице

\usepackage{tikz} % картинки в tikz
\usepackage{microtype} % свешивание пунктуации

\usepackage{verbatim} % для комментов

\usepackage{array} % для столбцов фиксированной ширины

\usepackage{indentfirst} % отступ в первом параграфе

\usepackage{sectsty} % для центрирования названий частей
\allsectionsfont{\centering}

\usepackage{amsmath} % куча стандартных математических плюшек
\usepackage{amssymb} % и символов
\usepackage{bbm}

\usepackage{multicol} % текст в несколько колонок

\usepackage{lastpage} % чтобы узнать номер последней страницы

\usepackage{enumitem} % дополнительные плюшки для списков
%  например \begin{enumerate}[resume] позволяет продолжить нумерацию в новом списке




\usepackage{fontspec} % хз
\usepackage{polyglossia} % для выбора языка в xelatex

\setmainlanguage{english}
\setotherlanguages{english}

% download "Linux Libertine" fonts:
% http://www.linuxlibertine.org/index.php?id=91&L=1
\setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
% why do we need \newfontfamily:
% http://tex.stackexchange.com/questions/91507/
\newfontfamily{\cyrillicfonttt}{Linux Libertine O}

\AddEnumerateCounter{\asbuk}{\russian@alph}{щ} % для списков с русскими буквами
%\setlist[enumerate, 2]{label=\asbuk*),ref=\asbuk*} % списки уровня 2 будут буквами а) б) ...

\usepackage{todonotes} % для вставки в документ заметок о том, что осталось сделать
% \todo[inline]{Здесь надо коэффициенты исправить}
% \missingfigure{Здесь будет картина Последний день Помпеи}
% команда \listoftodos — печатает все поставленные \todo'шки

\usepackage{booktabs} % красивые таблицы
% заповеди из документации:
% 1. Не используйте вертикальные линии
% 2. Не используйте двойные линии
% 3. Единицы измерения помещайте в шапку таблицы
% 4. Не сокращайте .1 вместо 0.1
% 5. Повторяющееся значение повторяйте, а не говорите "то же"


% \usepackage[left=1cm,right=1cm,top=1cm,bottom=1cm]{geometry}

\usepackage{fancyhdr} % весёлые колонтитулы
\pagestyle{fancy}
\lhead{Mathematics, exam}
\chead{}
\rhead{2017-12-29}
\lfoot{}
\cfoot{}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

\DeclareMathOperator{\E}{\mathbb{E}}
\let\P\relax
\DeclareMathOperator{\P}{\mathbb{P}}
\DeclareMathOperator{\Var}{\mathbb{V}ar}
\DeclareMathOperator{\Cov}{\mathbb{C}ov}



%% эконометрические сокращения
\def \hb{\hat{\beta}}
\DeclareMathOperator{\sVar}{sVar}
\DeclareMathOperator{\sCov}{sCov}
\DeclareMathOperator{\sCorr}{sCorr}

\def \1{\mathbbm{1}}

\def \hs{\hat{s}}
\def \hy{\hat{y}}
\def \hY{\hat{Y}}
\def \he{\hat{\varepsilon}}
\def \v1{\vec{1}}
\def \cN{\mathcal{N}}
\def \e{\varepsilon}
\def \z{z}

\def \hVar{\widehat{\Var}}
\def \hCorr{\widehat{\Corr}}
\def \hCov{\widehat{\Cov}}

\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator*{\plim}{plim}

%% лаг
\renewcommand{\L}{\mathrm{L}}


\begin{document}

\section{Dynamic Optimization}

\begin{enumerate}


\item {[10 points]} Consider a system of differential equations
\[
\begin{cases}
  \dot x = - 2y + x({x^2} + {y^2} - 1);\\
  \dot y =  2x + y({x^2} + {y^2} - 1).\\
\end{cases}
\]
\begin{enumerate}
  \item {[5 points]} Check that $(0,0)$ is the only point of rest of this system. Change the variables from the Cartesian to the polar $(r, \phi )$. Rewrite the system in the polar coordinates.
  \item {[5 points]} Draw two time paths of the system for $r_0 > 1$, where $r_0$ is the radius of the initial point and for $0 < r_0 < 1$. Justify your sketch. What if $r_0 = 1$?
\end{enumerate}

\item {[10 points]} Solve the calculus of variations problem \[
\int_1^2 \left( \frac{2y^2}{x^2} + (y')^2 \right) \, dx \to \max /\min
\]
with the fixed endpoints $y(1) = 0$, $y(2) = \frac{7}{2}$.

Hint: for solving Euler’s equation try to find solutions in the form of $x^a$. Check all available second-order conditions.

\item {[20 points]} Road construction costs minimization.

Let the terrain profile be represented by the function
\[
y(t) =
\begin{cases}
3 - 3|t|, \text{ if } |t|\leq 1 \\
0, \text{ otherwise.} \\
\end{cases}
\]

The contractor minimizes the excavation costs given by the formula
\[
\int_{-c}^c (x(t) - y(t))^2 \, dt,
\]
where $x(t)$ is the road profile we need to find, $c > 2$ and $[ - c,c]$ — road section where the excavation takes place ($c$ is not set). Allowable grade of the road satisfies  $| \dot x | \leq 1$.

\begin{enumerate}
%\item State the optimal bounded control problem with  constraint.
\item {[5 points]} Form Hamiltonian and derive first-order conditions taking into account that multiplier $\lambda$ satisfies transversality conditions $\lambda(-c) = \lambda(c) = 0$.
\item {[15 points]} Find $x(t)$ on the section $[ - c,c]$.
\end{enumerate}

\end{enumerate}



\newpage
\section{Stochastic Calculus}

Standard Wiener process is denoted by $W_t$.

\begin{enumerate}
  \item {[10 points]} Let $Y$ be equal to $1$ if $W_2>0$ and $0$ otherwise.

  Find $\E(Y|W_1)$, $\Var(Y|W_1)$, $\E(Y|W_1^2)$.
  \item {[10 points]} James Bond flips a biased coin until the sequence Head-Tail-Head-Tail appears. The probability of «Head» is equal to $p$. Using Doog's theorem and an appropriate martingale find the expected number of coin flips.
  \item {[10 points]} The process $X_t$ is given by
  \[
    X_t = 2017 + t^2 W_t^2 + \int_0^t u \, dW_u
  \]
  \begin{enumerate}
    \item Find $dX_t$;
    \item Is $X_t$ a martingale?
    \item Find $\E(X_t)$.
  \end{enumerate}

  \item {[10 points]} Consider the framework of the Black and Scholes model. You agreed with Warren Buffett that at fixed time $T$ he will pay you the strange sum
  \[
  X_T = \ln S_T \cdot \ln S_{T/2},
  \]
where $S_t$ is the price of a share.

What is the non-arbitrage price $X_0$ of this agreement?
  \item {[20 points]} Solve the stochastic differential equation
  \[
  dY_t =(Y_t^3 - Y_t) \, dt + Y_t^2 \, dW_t
  \]

  You are free to use or not to use the following guiding steps:

  \begin{enumerate}
    \item {[5 points]} Consider $Z_t = Y_t^n$ and find $dZ_t$;
    \item {[3 points]} Find a constant $n$ such that the term before $dW_t$ in $dZ_t$ is non-random.
    \item {[2 points]} Write down the equation for $dZ_t$ in terms of $t$ and $Z_t$ only: you should get rid of $Y_t$.
    \item {[8 points]} Solve the equation for $dZ_t$.

    Hint: It's just the particular case of the equation in your hometask~:) Do you remember? Multiply $Z_t$ by some exponent~:)
    \item {[2 points]} Finally, find $Y_t$.
  \end{enumerate}

\begin{comment}
  % Надо подумать ещё!!!!
  \item Solve the stochastic differential equation
  \[
  dX_t =(X_t^3 + X_t) \, dt + X_t \, dW_t
  \]

  You are free to use or not to use the following guiding steps:

  \begin{enumerate}
    \item Consider $Y_t = f(X_t)$ and find $dY_t$;
    \item State equation on $f$ such that in $dY_t$ the term before $dt$ is non-random.
    \item Find at least one
    \item Solve the equation for $Y_t$ and then find $X_t$;
  \end{enumerate}
\end{comment}

\end{enumerate}




\end{document}
