\documentclass[pdftex,12pt,a4paper]{article}

\input{/home/boris/science/tex_general/title_bor_utf8}

\emergencystretch=2em \voffset=-2cm \hoffset=-2cm
\unitlength=0.6mm \textwidth=18cm \textheight=26cm


\begin{document}
\parindent=0 pt % отступ равен 0

Stochastic calculus and dynamic optimization retake.

In all problems $W_t$ denotes the standard Wiener process.

\begin{enumerate}
\item 15 points. The random variables $X_1$, $X_2$, \ldots are independent uniformly distributed on $[0;1]$. I~am summing them until the first $X_i$ greater than $0.5$ is added. After this term I stop. Let's denote by $S$ the total sum and by $N$ --- the number of terms added. Find $\E(S|N)$, $Var(S|N)$.

Hint: to speed up the computations you may use the fact that $Var(X_i)=1/12$ without proof.
\item 15 points. Consider the process $X_t=\int_0^t s^2W_s\,dW_s$. Find $\E(X_t)$, $Var(X_t)$, $Cov(X_t,W_t)$

\item 20 points. For $0<s<t$ calculate the probability $\P(W_t>0 \mid W_s>0)$. 
\item 25 points. Consider the following stochastic differential equation
\begin{equation}
dX_t=(\sqrt{1-X_t^2}-0.5X_t)dt+\sqrt{1-X_t^2}dW_t, \quad X_0=0
\end{equation}
\begin{enumerate}
\item Consider a substitution $X_t=\sin Y_t$, where $Y_t$ is some Ito process $dY_t=A_tdt+B_tdW_t$. Using Ito's lemma find $dX_t$.
\item State conditions on $A_t$ and $B_t$ such that the two $dX_t$ coincides
\item Find $Y_t$ and $X_t$
\end{enumerate}
\item 25 points. Consider the profit-maximizing problem for a representative competitive firm 
\begin{equation}
\int_0^{\infty}(p-c(x(t)))q(t)e^{-rt}\,dt\rightarrow \max
\end{equation}
 where the state variable $x(t)<1$ represents a nonrenewable stock resource (oil) that depletes according to the equation $\dot{x}=1-x-q$ and $q(t)$ is the extraction rate. Here $c(x)$ is a cost function of the extraction that is defined by $c(x)=e^{-x}$. The price of oil is assumed to be constant and equal $p$ where $e^{-1}<p<1$. The optimization problem is to choose $q(t)$ to maximize the discounted profits, $0<r<1$.
\begin{enumerate}
\item Derive necessary conditions.
\item Prove that the steady-state exists and is unique.
\end{enumerate}
\end{enumerate}


\end{document}