\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[russian]{babel}


\usepackage{url}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[left = 1cm, right = 1cm, top = 2cm, bottom = 2cm]{geometry}

\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{Graded hometask on stochastic calculus}
\chead{}
\rhead{ICEF. Fall 2015}


\DeclareMathOperator{\Cov}{Cov}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\E}{E}



\begin{document}

You should solve five out of six problems at your choice.

\begin{enumerate}

\item Find the expected value of $\E(\exp(aW_t))$, $\E(1_{W_t\leq b})$, $\E(\exp(aW_t)\cdot 1_{W_t\leq b})$ and $\E(W_t 1_{W_t\leq b})$. Naturally, you may use the standard normal cumulative distribution function $F$ in your answer


\item It is known that $\E(Y|X)=0$. Which of the following quantities must be zero: $\E(Y)$? $\E(X)$? $\Cov(X,Y)$? $\Cov(X^2,Y)$? $\Cov(X,Y^2)$? Prove or provide a counter-example.

\item Alisa and Bob throw a fair coin until either the sequence THTH or the sequence HTHH appears. Alisa wins if the sequence THTH appears first and Bob wins if the sequence HTHH appears first.
\begin{enumerate}
\item What is the probability that Alisa will win?
\item What is the expected duration of the game in tosses?
\end{enumerate}

Hint: you may introduce a martingale from lecture or solve this problem without martingales at all

\item Consider the process $Y_t=\exp(2W_t-2t)$.

\begin{enumerate}
\item Find $dY_t$
\item Find $\int_0^t Y_u \, dW_u$
\item Find $\E(Y_t)$ and $\Var(Y_t)$
\end{enumerate}

\item Consider stochastic differential equation
\[
dX_t = (a - bX_t) dt + c dW_t, \; X_0=x_0
\]

\begin{enumerate}
\item Solve this differential equation\footnote{The answer may contain an Ito integral that cannot be simplified.}.
\item Find $\E(X_t)$ and $\Var(X_t)$
\end{enumerate}

\item In the framework of Black and Scholes model find the price at $t=0$ of the classic European Ñall option by calculating corresponding expected value.

European call option with strike price $K$ is the right to buy at time $t$ one share at price $K$. So, at time $t$ it pays you $S_t - K$ if $S_t > K$ and zero otherwise.

\end{enumerate}

\vspace{2cm}

You may decide not to solve any number of problems from this hometask and send me a solution of corresponding number of problems from old exams collection instead. In this case you should use \LaTeX{} or markdown formats. Please, be polite and avoid sending solutions of already solved problems or problems taken by someone else. You may choose problems and fill in your choice at \url{http://goo.gl/8TFhWE}.

The due date for this hometask --- 11.01.2016. If you decide to solve problems from an old exam you should send the solution of these problems by e-mail to \href{mailto:boris.demeshev@gmail.com}{boris.demeshev@gmail.com} before 09.01.2016. I will check your solutions and add them into the exams collection so they will be available for everyone. You will receive extra bonus points in this case.

\end{document}
